\vskip-0.6in Endogenous Politics and the Design of Trade Institutions

### 2015-01-08 ex_ante.tex
convert lines 69-96 ex_ante to text

Thus I model the influence of ex-ante lobbying on the governments' tariff-setting behavior as operating through a threshold. Recall how the trade agreement tariff is determined under ex-post lobbying only. Because the trade agreement acts as a commitment device, the governments choose the tariff level $\tau^R_{W,e}$, and implicitly $e^{EP}=e^R_{W,e}$, to maximize their joint political welfare as in Equation~\ref{exp:1}, subject only to the constraint that $\ga$ will be realized according to the ex-post lobbying process.

As discussed in Section~\ref{sec:objfcn}, depending on how government welfare changes as political pressure increases, the governments may be looking ahead and manipulating ex-post lobbying incentives when there is only ex-post lobbying. When ex-ante lobbying is possible the governments may have to raise the trade agreement tariff above $\tau^R_{W,e}$ in order to respond to that pressure.

This is modeled here with a simple cut-off. The trade agreement tariff cap will be set according to $\ga\left(e^{EA}\right)$ instead of at $\tau^R_{W,e}$ if $\ga\left(e^{EA}\right)$ would lead to a higher trade agreement tariff. If, on the other hand, $\tau = \frac{4(\ga(e^{EA})-1)}{25-4\ga(e^{EA})} < \tau^R_{W,e}$, the governments respond to the higher ex-post pressure and set $\tau^R_{W,e}$.

Thus the governments lose some measure of control with ex-ante lobbying. If $e^{EA}$ is low enough, ex-ante lobbying has no impact. But if $e^{EA}$ is above the threshold, the trade agreement tariff is strictly greater than $\tau^R_{W,e}$ and so joint government welfare will not be at its globally optimal level. The trade agreement the governments choose will be constrained by the ex-ante pressure.

@@

\subsection{Weak Bindings}
\begin{itemize}
	\item Here, $e^{EP}$ will also work as in body of paper. The earlier stage is sunk, and applied level will match whatever effort lobby puts forth
		\begin{itemize}
			\item It's just that cap could be higher due to ex-ante lobbying
				\begin{itemize}
					\item Without ex-ante lobbying, government chooses $\tau^R_{W,e}$ that leads to ex-post $e$ that maximizes ex-ante joint welfare
		\[
		  W(\ga(e),\tau^R_{W,e}) + W^*(\tau^R_{W,e})
		\]
		knowing they'll set $\tau^N$ in the third stage (i.e. $\tau^N(\ga(e^R_{W,e})) = \frac{8\ga(e^R_{W,e}) - 5}{68 - 8 \ga(e^R_{W,e})} = \tau^R_{W,e}$).
						\begin{itemize}
							\item $\tau^R_{W,e}$ is the optimal unilateral choice of the home government when the lobby is faced with the constraint of the trade agreement
							\item Lobby would max $\pi(\tau(\ga(e))) -e$; instead sets $e = e^R_{W,e} \Rightarrow \tau^N(\ga(e^R_{W,e})) =\tau^R_{W,e}$
						\end{itemize}
					\item Government actually ends up with a lower $e$ mapping into a lower $\ga$ to produce the trade agreement $\tau^R_{W,e}$ according to $\tau^N$, so has to set $\tau^R_{W,e}$ higher to target preferred $\ga$ according to joint max problem [in this symmetric case, identical to unilateral problem]
			\end{itemize}
		\end{itemize}
	\item Lobby decides $e^{EA}$ given that gov't will set cap $\tau^R_{W,e}$ or in ex-ante stage given $e^{EA}$, whichever is higher
		\begin{itemize}
			\item Start with calculus condition
				\[
				  \pi(\tau^E(e^{EA})) - e^{EA} - e^{EP}(e^{EA})
				\]
				When this FOC is operative, government decides tariff cap according to joint decision problem with $e^{EA}$, but then lobby still has to exert effort up to that cap; $e^{EP}$ is the amount the lobby pays according to $\tau^N$ to achieve $\tau^E(e^{EA})$
					\begin{itemize}
						\item FOC
							\begin{equation}
							  \frac{\partial \pi}{\partial \tau} \frac{\partial \tau^E}{\partial \ga} \frac{\partial \ga}{\partial e^{EA}} = 1 +\frac{\partial e^{EP}}{\partial e^{EA}}
								\label{eq:ea}
							\end{equation}
					\end{itemize}
			\item Then, set to zero if
			  \begin{equation}
				  \pi(\tau^E(e^{EA})) - e^{EA} - e^{EP}(e^{EA}) < \pi(\tau^N(e^R_{W,e})) - e^R_{W,e} 
					\label{ine:ea}
				\end{equation}
			\item Intuition: If lobby's incentives to exert pressure ex-ante are strong enough, they overpower the government's ability to use the trade agreementto directly manipulate lobbying incentives ex-post
				\begin{itemize}
					\item Lobby can force government to set cap higher than the level that would optimize its welfare when there is no ex-ante lobbying
					\item Result: With tariff cap, ex-ante lobbying can only (weakly) reduce gov't political welfare. It can only (weakly) increase net profits for lobby (lobby can set $e^{EA}=0$ if ex-ante lobbying would make it worse off)
					\item i.e. tariff with ex-ante lobbying can only be (weakly) higher than without ex-ante lobbying
				\end{itemize}	
		\end{itemize}	
\end{itemize}

%\vskip.5in

@@

\newpage
\vskip.2in
Compare to ex-ante lobbying \`{a} la MRC.
\begin{itemize}
	\item Cooperative bargaining soln (maximize joint welfare of both governments and both lobbies) instead of extensive form
		\begin{itemize}
			\item leads to bizarre outcome where government will take into account lobby's profits even if $e^{\text{EA}}=0$.
		\end{itemize}
	\item In MRC main treatment, assume lobby has all the bargaining power
		\begin{itemize}
			\item Implies tariff caps and strong bindings equivalent, even though these split the surplus in very different ways. All surplus goes to lobby
			\item In the part that is only in the NBER working paper, they have
				\begin{itemize}
					\item $z=0$ (perfectly immobile capital)
					\item no ex-ante lobbying
					\item government has all the bargaining power
					\item government sometimes chooses free trade---so there IS an incentive to use the trade agreement as a domestic commitment device
				\end{itemize}
			\end{itemize}
	\item In MRC, no matter how tariff is set (cooperatively or non-cooperatively), contribution is $xp(\ov{t})$
		\begin{itemize}
			\item Not true in GH95
			\item In my set-up, lobby's incentives change when tariff-setting process changes
		\end{itemize}
	\item In MRC, weight on lobby's profits is $1+a$ no matter what. Doesn't vary in lobby's efforts. GH not really microfoundations for flexible model as in Long and Vousden (1991) / Baldwin (1987) model; microfoundations for more restrictive version with fixed weights; then decide $\tau$ to change PS, CS, etc. (p. 481 GH94). Can't get shocks as in this literature
		\begin{itemize}
			\item I still don't know exactly why you never want to reign in lobbying in GH; because it's just a conduit for information, and gov't is fully compensated for loss? But is it?
		\end{itemize}
	\item Bargaining power
		\begin{itemize}
			\item It's important in MRC's story: they show how results change if they change bargaining power, and it's dramatic b/c they take an extreme view that lobbies have all the power
			\item I view bargaining power in this environment as an artificial construct---what's going on isn't really bargaining; bargaining is used as a reduced form. But it's a much more complex process
		\end{itemize}
\end{itemize}

>>>>

Thus I model the influence of ex-ante lobbying on the governments' tariff-setting behavior as operating through a threshold. Recall how the trade agreement tariff is determined under ex-post lobbying only. Because the trade agreement acts as a commitment device, the governments choose the tariff level $\tau^R_{\cdot,e}$, and implicitly $e^{EP}=e^R_{\cdot,e}$, to maximize their joint political welfare as in Equation~\ref{exp:1}, subject only to the constraint that $\ga$ will be realized according to the ex-post lobbying process.

As discussed in Section~\ref{sec:objfcn}, depending on how government welfare changes as political pressure increases, the governments may be looking ahead and manipulating ex-post lobbying incentives when there is only ex-post lobbying. When ex-ante lobbying is possible the governments may have to raise the trade agreement tariff above $\tau^R_{\cdot,e}$ in order to respond to that pressure.

This is modeled here with a simple cut-off. The trade agreement tariff cap will be set according to $\ga\left(e^{EA}\right)$ instead of at $\tau^R_{\cdot,e}$ if $\ga\left(e^{EA}\right)$ would lead to a higher trade agreement tariff. If, on the other hand, $\tau = \frac{4(\ga(e^{EA})-1)}{25-4\ga(e^{EA})} < \tau^R_{\cdot,e}$, the governments respond to the higher ex-post pressure and set $\tau^R_{\cdot,e}$.

Thus the governments lose some measure of control with ex-ante lobbying. If $e^{EA}$ is low enough, ex-ante lobbying has no impact. But if $e^{EA}$ is above the threshold, the trade agreement tariff is strictly greater than $\tau^R_{\cdot,e}$ and so joint government welfare will not be at its globally optimal level. The trade agreement the governments choose will be constrained by the ex-ante pressure.

@@

\subsection{Weak Bindings}
In the case of weak bindings, or tariff caps, the ex-post stage will also work as in the no-ex-ante-lobbying case, that is, as in Section~\ref{sec:weak}, again because ex-ante effort is sunk at the ex-post stage. To recap, when there is no ex-ante lobbying---or ex-ante-lobbying is less important than ex-post lobbying---the governments choose the $\tau^R_{W,e}$ that leads to the level of ex-post political pressure that maximizes ex-ante joint welfare
		\[
		  W(\ga(e^R_{W,e})),\tau^R_{W,e}) + W^*(\tau^R_{W,e})
		\]
		knowing that the home government will set the applied tariff in the third stage unilaterally (i.e. $\tau^N(\ga(e^R_{W,e})) = \frac{8\ga(e^R_{W,e}) - 5}{68 - 8 \ga(e^R_{W,e})} = \tau^R_{W,e}$).

What is different here is that the tariff cap could be set higher than this level because of the presence of ex-ante lobbying. The lobby chooses its ex-ante effort level $e^{EA}$ taking into account that the governments will set either the optimal ex-post tariff cap $\tau^R_{W,e}$ or a higher cap in ex-ante stage given the $e^{EA}$ with which it is confronted.

The lobby maximimizes $\pi(\tau^E(e^{EA})) - e^{EA} - e^{EP}(e^{EA})$, with a first order condition of 
	\begin{equation}
	  \frac{\partial \pi}{\partial \tau} \frac{\partial \tau^E}{\partial \ga} \frac{\partial \ga}{\partial e^{EA}} = 1 +\frac{\partial e^{EP}}{\partial e^{EA}}
			\label{eq:ea}
	\end{equation}
This condition illustrates that, when it is operative, the governments choose the tariff cap according to Expression~\ref{exp:1} with $e^{EA}$, but then lobby still has to exert effort ex-post to receive the applied tariff. $e^{EP}$ is the amount the lobby pays according to Equation~\ref{eq:nash} to achieve $\tau^E(e^{EA})$ (see Equation~\ref{eq:eff}).

The lobby sets $e^{EA}$ according to this condition as long as 
	\begin{equation}
		\pi(\tau^E(e^{EA})) - e^{EA} - e^{EP}(e^{EA}) > \pi(\tau^N(e^R_{W,e})) - e^R_{W,e}.
		\label{ine:ea}
	\end{equation}
Otherwise, it maximizes net profits by setting $e^{EA} =0$.

Thus if the lobby's incentives to exert pressure ex-ante are strong enough, they overpower the government's ability to use the trade agreement to directly manipulate lobbying incentives ex-post. The lobby can force the government to set a cap higher than the level that would optimize its welfare when there is no ex-ante lobbying. It has no interest in a lower cap and can choose a zero effort level whenever positive effort would reduce its nets profits below the level with ex-post lobbying only. Therefore we have the following reesult:

\begin{result}
	When trade agreements take the form of weak bindings, the binding under ex-ante lobbying can only be (weakly) higher than when ex-ante lobbying is not possible. This can only (weakly) reduce the joint political welfare of the governments and it can only (weakly) increase net profits for the lobby.
\end{result}

@@

\newpage
\vskip.2in
Compare to ex-ante lobbying \`{a} la MRC.
\begin{itemize}
	\item Cooperative bargaining soln (maximize joint welfare of both governments and both lobbies) instead of extensive form
		\begin{itemize}
			\item leads to bizarre outcome where government will take into account lobby's profits even if $e^{\text{EA}}=0$.
		\end{itemize}
	\item In MRC main treatment, assume lobby has all the bargaining power
		\begin{itemize}
			\item Implies tariff caps and strong bindings equivalent, even though these split the surplus in very different ways. All surplus goes to lobby
			\item In the part that is only in the NBER working paper, they have
				\begin{itemize}
					\item $z=0$ (perfectly immobile capital)
					\item no ex-ante lobbying
					\item government has all the bargaining power
					\item government sometimes chooses free trade---so there IS an incentive to use the trade agreement as a domestic commitment device
				\end{itemize}
			\end{itemize}
	\item In MRC, no matter how tariff is set (cooperatively or non-cooperatively), contribution is $xp(\ov{t})$
		\begin{itemize}
			\item Not true in GH95
			\item In my set-up, lobby's incentives change when tariff-setting process changes
		\end{itemize}
	\item In MRC, weight on lobby's profits is $1+a$ no matter what. Doesn't vary in lobby's efforts. GH not really microfoundations for flexible model as in Long and Vousden (1991) / Baldwin (1987) model; microfoundations for more restrictive version with fixed weights; then decide $\tau$ to change PS, CS, etc. (p. 481 GH94). Can't get shocks as in this literature
		\begin{itemize}
			\item I still don't know exactly why you never want to reign in lobbying in GH; because it's just a conduit for information, and gov't is fully compensated for loss? But is it?
		\end{itemize}
\end{itemize}
